On The Impact of Atoms of Confusion in JavaScript Code

### 2022-03-29 3_Method.tex
Added text describing the methodology of the replication study as a separate file. This commit aims to be an intermediate step. The next one will be to combine the methodologies of both studies, in particular the two experiments.

We decided not to provide feedback about the time students took to answer each question. We also did not tell them whether their answers were correct or not. Our main concern was to avoid introducing bias for future respondents. Since we posted the survey in a social media platform, possible threats could have arisen if we gave respondents instant feedback.

@@

\subsubsection*{Interview Process} We conducted semi-structured interviews using a web conferencing software. We recorded all the interviews with the consent of the participants. On average, the interviews lasted 26.29 minutes, with the shortest one lasting 14.59 minutes and the longest one 43.06 minutes. 
Two of the researchers conducted the interviews, and a third one listened to all the recordings to cross-validate the collected data.
The interviews had three main parts. In the first one, we asked the developers the following demographic information: name, email, gender, level of education, current job position, JavaScript experience in years and other programming languages they have worked with.
Table \ref{pinterview} summarizes this demographic information.

@@

We mined the occurrence of atom candidates from the repositories in our curated dataset using source code queries that we wrote using the CodeQL language~\cite{moor:gttse2007}. CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as \cpplang, Java, and JavaScript). We also automate the process of running the queries and exporting the results to a format that simplifies our analysis (and also the reproduction of this study). Finally, we carried out basic descriptive statistics in order to understand how often the atoms of confusion appear in practice.

>>>>

We decided not to provide feedback about the time students took to answer each question. We also did not tell them whether their answers were correct or not. Our main concern was to avoid introducing bias for future respondents. Since we posted the survey on a social media platform, possible threats could have arisen if we gave respondents instant feedback.

@@

\subsubsection*{Interview Process} We conducted semi-structured interviews using web conferencing software. We recorded all the interviews with the consent of the participants. On average, the interviews lasted 26.29 minutes, with the shortest one lasting 14.59 minutes and the longest one 43.06 minutes. 
Two of the researchers conducted the interviews, and a third one listened to all the recordings to cross-validate the collected data.
The interviews had three main parts. In the first one, we asked the developers the following demographic information: name, email, gender, level of education, current job position, JavaScript experience in years, and other programming languages they have worked with.
Table \ref{pinterview} summarizes this demographic information.

@@

We mined the occurrence of atom candidates from the repositories in our curated dataset using source code queries that we wrote using the CodeQL language~\cite{moor:gttse2007}. CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as \cpplang, Java, and JavaScript). We also automate the process of running the queries and exporting the results to a format that simplifies our analysis (and also the reproduction of this study). Finally, we computed some descriptive statistics to measure the prevalence of atoms of confusion in practice.

### 2022-03-29 6_Conclusion.tex
Added text describing the methodology of the replication study as a separate file. This commit aims to be an intermediate step. The next one will be to combine the methodologies of both studies, in particular the two experiments.

This paper reports the results of a mixed-method research effort to better understand the impact of atoms of confusion in JavaScript code. First, we conducted a survey with 140 JavaScript developers, asking them to predict the output of some code snippets (with and without atoms). We provide evidence that five atom candidates significantly impact program comprehension activities and can be considered atoms of confusion: Comma Operator, Automatic Semicolon Insertion, Post Increment, Assignment as Value, and Implicit Predicate. 
We also interviewed 15 professional developers. In general, they considered code without atom candidates easier to understand and highlighted other JavaScript constructs and idioms that might also introduce misunderstanding. 
%Last, but not least,
Finally, we conducted a mining software repositories study. We found four atom candidates frequently used in JavaScript code, i.e., Implicit Predicate, Post Increment, Ternary Operator, and Omitted Curly Braces.  %The first two can be considered atoms of confusion.

Our efforts have important implications for practice. First, we alert developers to avoid writing JavaScript code with certain atoms of confusion (e.g., Comma Operator, Automatic Semicolon Insertion, Post Increment, Assignment as Value, and Implicit Predicate). Second, our results might help tool developers to create program transformation tools to remove the atoms automatically. 
%They might firstly focus, for instance, on the atoms we found to be more confusing and at the same time more common in practice (e.g., Omitted Curly Braces and Post Increment).
%Promising candidates may be the atoms we found to be most confusing and, at the same time, common in practice (e.g., Implicit Predicate and Post Increment).

%As mentioned, the results of the interviews
The results of the interviews pointed out additional JavaScript constructs and idioms 
that might introduce misunderstanding. These include nested callbacks, % and prototype-based inheritance (which might not have a counterpart version), but also idioms such
%as 
\emph{property access}, and \emph{arrow functions}. As future work, we intend to
reproduce our survey to validate whether or not these additional
\emph{atom candidates}
%, which the interviewees pointed out, 
truly affect
the understanding of JavaScript code.

>>>>

This paper reports the results of a mixed-method research effort that allowed us to better understand the impact of atoms of confusion in JavaScript code. First, we conducted a survey with 140 JavaScript developers, asking them to predict the output of some code snippets (with and without atoms). We provide evidence that five atom candidates significantly impact program comprehension activities and can be considered atoms of confusion.
%% : Comma Operator, Automatic Semicolon Insertion, Post Increment, Assignment as Value, and Implicit Predicate. 
%% We also interviewed 15 professional developers. In general, they considered code without atom candidates easier to understand and highlighted other JavaScript constructs and idioms that might also introduce misunderstanding. 
%Last, but not least,
%Finally, we conducted a mining software repositories study. We found four atom candidates frequently used in JavaScript code, i.e., Implicit Predicate, Post Increment, Ternary Operator, and Omitted Curly Braces.  %The first two can be considered atoms of confusion. 
Our efforts have two implications for practice. First, we present evidence that not all atoms of confusion validated to statically typed languages (such as \clang, \cpplang, or Java) led to a statistically significant impact on program understanding for JavaScript (a dynamically typed language). Besides that, our findings might be used to alert developers to avoid writing JavaScript code with certain atoms of confusion (e.g., Comma Operator, Automatic Semicolon Insertion, Post Increment/Decrement, Assignment as Value, and Implicit Predicate). Finally, our results might help tool developers to create program transformation tools for removing atoms that frequently appear in software. 
%They might firstly focus, for instance, on the atoms we found to be more confusing and at the same time more common in practice (e.g., Omitted Curly Braces and Post Increment).
%Promising candidates may be the atoms we found to be most confusing and, at the same time, common in practice (e.g., Implicit Predicate and Post Increment).
%As mentioned, the results of the interviews
Our research also pointed out additional JavaScript constructs and idioms 
that might introduce misunderstanding. These include nested callbacks, % and prototype-based inheritance (which might not have a counterpart version), but also idioms such
%as 
\emph{property access}, and \emph{arrow functions}. As future work, we intend to
reproduce our survey to validate whether or not these additional
\emph{atom candidates}
%, which the interviewees pointed out, 
truly affect
the understanding of JavaScript code.
